---
title: "Assignment_3_3102_bv10"
author: Bhuvaneswari(Bhu) Venkatraman
netid: bv10
seedid: 3102
output:
  html_document: 
    theme: readable
    toc: yes
    toc_float: true
---

# Set seed

```{r}
library(splines)
set.seed(3102)
```

# PART - 1
## Prepare your function

## Write your own function to use LOO-CV and GCV to select the optimal span for loess.

```{r}
lo.lev <- function(x1, sp){
  # x1: n-by-1 feature vector
  # sp: a numerical value for "span"
  
  n = length(x1);
  lev = rep(0, n)
  
  ##############################################
  # YOUR CODE: Compute the diagonal entries of the 
  #            smoother matrix S and 
  #            store it in a vector "lev"
  # Tip: check how we compute the smoother matrix
  #      for smoothing spline models
  
  for(i in 1:n){
    y = rep(0, n); y[i]=1;
    yi = loess(y ~ x1, span = sp)$fitted;
    lev[i]= yi[i];
  }
  
  ##############################################
  
  return(lev)
}

onestep_CV <- function(x1, y1, sp){
  
  ##############################################
  #  YOUR CODE: 
  #  1) Fit a loess model y1 ~ x1 with span = sp, and extract 
  #     the corresponding residual vector
  #  2) Call lo.lev to obtain the diagonal entries of S
  #  3) Compute LOO-CV and GCV using formula from lecture notes
  #    [lec_W5_NonlinearRegression.pdf] page 33. 
  
  n = length(x1)
  err = y1 - loess(y1 ~ x1, span = sp)$fitted
  sdiag = lo.lev(x1, sp)
  
  cv = sum((err/(1-sdiag))^2)/n
  gcv = sum((err/(1-sum(sdiag)/n))^2)/n
  ##############################################
  
  return(list(cv = cv, gcv = gcv))
}

myCV <- function(x1, y1, span){
  # x1: feature vector of length n
  # y1: response vector of length n
  # span: a sequence of values for "span"
  
  m = length(span)
  cv = rep(0, m)
  gcv = rep(0, m)
  
  for(i in 1:m){
    tmp = onestep_CV(x1, y1, span[i])
    cv[i] = tmp$cv
    gcv[i] = tmp$gcv
  }
  return(list(cv = cv, gcv = gcv))
}
```

### Test your function

```{r}
mydata = read.csv(file = "Coding3_Data.csv")
dim(mydata)
plot(mydata$x, mydata$y, xlab="", ylab="")
```


### Create a grid of values for span: 15 values that are equally spaced between 0.20 and 0.90. Call your function myCV to compute the corresponding LOO-CV and GCV.

```{r}
span1 = seq(from = 0.2, by = 0.05, length = 15 )
cv.out = myCV(mydata$x, mydata$y, span1)
```

## Print out your results
```{r}
myout = data.frame(CV = cv.out$cv, 
                   GCV = cv.out$gcv, 
                   span = span1)
myout
```
```{r}
myout$span[myout$GCV == min(myout$GCV)]
```

```{r}
myout$span[myout$CV == min(myout$CV)]
```

## Plot the fitted curve
### Plot the data (red circles), the true curve (gray) and the fitted curve (blue dashed line) using the optimal span.

```{r}
spangcv.min = 0.5
plot(mydata$x, mydata$y, xlab="", ylab="", col="gray");
fx = 1:50/50;
fy = sin(12*(fx+0.2))/(fx+0.2)
lines(fx, fy, col=8, lwd=2);
f = loess(y ~ x, mydata, span = spangcv.min)
lines(fx, predict(f, data.frame(x = fx), surface = "direct"), 
      lty=2, lwd=2, col="blue")
```

# PART - 2

## Load Data
```{r}
set.seed(3102) 
mydata = read.csv("Sales_Transactions_Dataset_Weekly.csv")
ts = as.matrix(mydata[, 2:53])
row.names(ts) = mydata[,1]
ts = ts - rowMeans(ts)
```

## Calculate B value

```{r}

x = seq(0, 1, length.out = ncol(ts))
F = ns(x, df = 10, intercept = FALSE)
F = t(t(F) - colMeans(F))
B = solve(t(F) %*% F) %*% t(F) %*% t(ts)


```

## Clustering with B
### Run k-means algorithm on the coefficient matrix B to cluster the 811 products into 6 clusters. Display time series for products in the same cluster in one figure along with the corresponding cluster center; arrange the 6 figures in 2-by-3 format.


```{r}
mykm1 = kmeans(t(B), 6)
myK = 6
mycenters1=mykm1$centers
mycenters1_6clusters=F %*% t(mycenters1)
par(mfrow=c(2,3))
for(k in 1:myK){
  id=which(mykm1$cluster==k)
  plot(NA, xlim = c(1, ncol(ts)), ylim = range(ts), 
       xlab = "Weeks", ylab = "Weekly Sales")
  for(i in 1:length(id))
    lines(1:ncol(ts), ts[id[i],] , col="gray")
  lines(1:ncol(ts), mycenters1_6clusters[,k], col="red")
}
```

## Clustering with X
### Run k-means algorithm on the original time series matrix X to cluster the 811 products into 6 clusters. Display time series for products in the same cluster in one figure along with the corresponding cluster center; arrange the 6 figures in 2-by-3 format.

```{r}
mykm2 = kmeans(ts, 6)
mycenters2 = t(mykm2$centers)
myK = 6
par(mfrow=c(2,3))
for(k in 1:myK){
  id=which(mykm2$cluster==k)
  plot(NA, xlim = c(1, ncol(ts)), ylim = range(ts), 
       xlab = "Weeks", ylab = "Weekly Sales")
  for(i in 1:length(id))
    lines(1:ncol(ts), ts[id[i],] , col="gray")
  lines(1:ncol(ts), mycenters2[,k], col="red")
}
```

